Temporal ordering and registration of images in studies of developmental dynamics

### 2015-02-02 image_analysis_paper_development2/manuscript.tex
updated report

\subsubsection*{Fixed {\em Drosophila} images}

@@

\subsubsection*{Wing disc data}

@@

Already at this point, the rate of collection of imaging data has surpassed the rate of manual image analysis.
%
This necessitates the development of automated methodologies to organize such large data sets. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have been explored largely independently of each other. 
%
Temporal ordering of large-scale data which did not require registration was done in the context of molecular profiling studies, in which data are vectors describing the expression levels of different mRNA \citep{anavy2014blind, trapnell2014dynamics, gupta2008extracting}.
%
At the same time, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point \citep{yuan2014automated, surkova2008characterization}.  
%
%The task of image registration has been widely studied \citep{zitova2003image}, for applications such as face recognition \citep{rowley1998rotation}, medical image registration \citep{hajnal2010medical}, and texture classification \citep{greenspan1994rotation}.
%
In contrast to most of the existing registration approaches which rely on the knowledge of appropriate landmarks in the images \citep{ian1998statistical} (such as the eyes in face recognition applications \citep{zhao2003face}), algorithms based on angular synchronization can register images even in the absence of such information, making them relevant for a wide variety of applications. 
%
We have applied our methodology to several different proprotypical data sets in developmental biology; because our methodology only relies on computing distances between images,, we are confident it may be readily extended to many other data sets.

We acknowledge that our methods, though general, do have some limitations. 
%
The first is that we require enough amount of data to sufficiently sample the developmental trajectory.
%
Therefore, for very small and/or very noisy data sets, our algorithms may fail. 
%
Second, the relevant developmental patterns and structures need to be large compared to the noise and the image resolution. 
%
In all of our examples, the relevant features, such as expression patterns and morphological structures, all span several pixels and are large compared to both the instrument noise and embryo-to-embryo variability. 
%
This means that the Euclidean distance is a good measure of similarity between images.
%
Our algorithms will fail for other imaging data sets which contain very fine scale structures,  as small perturbations in these fine-scale structures will result in large changes in the Euclidean distances. 
%
In such cases, one requires a more sophisticated distance metric, such as the Earth Mover's Distance \citep{...}, or image description, such as scattering transform coefficients \citep{...}, which are insensitive to such perturbations. 
%
Furthermore, as we have acknowledged, our methods as implemented cannot register full three-dimensional data with respect to rotations due to computational limitations. 
%
However, the algorithms presented here could be used for three-dimensional given an alternative method for computing pairwise alignments more efficiently (such as using image features or landmarks).

@@

\subsection*{{\em Drosophila} Experiments}
%
Oregon-R was used as wild type {\em Drosophila} strains. 
%
Embryos were collected and fixed at 22$^\circ$C. 
%
Monoclonal rabbit anti-dpERK (1:100, Cell signaling) and rat anti-Twist (1:500, a gift from Eric Wieschaus) were used to stain proteins of interest. 
%
DAPI (1:10,000, Vector Laboratories) was used to visualize nuclei, and Alexa Fluors (1:500, Invitrogen) were used as secondary antibodies. 
%
Histone-RFP strain was used to obtain time-lapse movie of gastrulating embryos at 22$^\circ$C. 
%
Embryos were loaded to the microfluidic device with PBST to keep them oxidized.

\subsection*{{\em Drosophila} Microscopy}
%
Nikon A1-RS scanning confocal microscope, and the Nikon 60x Plan-Apo oil objective was used to image embryos. 
%
Embryos were collected, stained, and imaged together under the same microscope setting. 
%
End-on imaging was performed by using the microfluidics device described previously \citep{chung2010microfluidic}.
%
Images were collected at the focal plane $\sim$90~$\mathrm{\mu m}$ from the posterior pole of an embryo (see \fig~\ref{fig:DV_view}).

@@

\subsection*{Software}
%
All algorithms and analysis were implemented in MATLAB\textsuperscript{\textregistered} (R2013b, The MathWorks, Natick, Massachusetts).
%
Code is available at \texttt{genomics.princeton.edu/stas/publications.html} under ``Codes and Data''.

>>>>

\subsubsection*{Fixed images of {\em Drosophila} gastrulation}

@@

\subsubsection*{Fixed z-stacks of {\em Drosophila} wing discs}

@@

Already at this point, the rate of collection of imaging data has surpassed the rate of manual image analysis.
%
This necessitates the development of automated methodologies to organize such large data sets. 
%
To the best of our knowledge, algorithmic approaches to these two tasks have been explored largely independently of each other. 
%
Temporal ordering of large-scale data which did not require registration was done in the context of molecular profiling studies, in which data are vectors describing the expression levels of different mRNA \citep{anavy2014blind, trapnell2014dynamics, gupta2008extracting}.
%
At the same time, temporal ordering of imaging data sets was done with a significant amount of human supervision and using registered images as a starting point \citep{yuan2014automated, surkova2008characterization}, or using some {\em a priori} knowledge of the relevant developmental processes \citep{dubuis2013accurate}.  
%
%The task of image registration has been widely studied \citep{zitova2003image}, for applications such as face recognition \citep{rowley1998rotation}, medical image registration \citep{hajnal2010medical}, and texture classification \citep{greenspan1994rotation}.
%
In contrast to most of the existing registration approaches which rely on the knowledge of appropriate landmarks in the images \citep{ian1998statistical} (such as the eyes in face recognition applications \citep{zhao2003face}), algorithms based on angular synchronization can register images even in the absence of such information, making them relevant for a wide variety of applications. 
%
We have applied our methodology to several different proprotypical data sets in developmental biology; because our methodology only relies on computing distances between images,, we are confident it may be readily extended to many other data sets.

We acknowledge that our methods, though general, do have some limitations. 
%
The first is that we require enough amount of data to sufficiently sample the developmental trajectory.
%
Therefore, for very small and/or very noisy data sets, our algorithms may fail. 
%
Second, the relevant developmental patterns and structures need to be large compared to the noise and the image resolution. 
%
In all of our examples, the relevant features, such as expression patterns and morphological structures, all span several pixels and are large compared to both the instrument noise and embryo-to-embryo variability. 
%
This means that the Euclidean distance is a good measure of similarity between images.
%
Our algorithms will fail for other imaging data sets which contain very fine scale structures,  as small perturbations in these fine-scale structures will result in large changes in the Euclidean distances. 
%
In such cases, one requires a more sophisticated distance metric, such as the Earth Mover's Distance \citep{rubner2000earth}, or image description, such as scattering transform coefficients \citep{mallat2012group}, which are insensitive to such perturbations. 
%
Furthermore, as we have acknowledged, our methods as implemented cannot register full three-dimensional data with respect to rotations due to computational limitations. 
%
However, the algorithms presented here could be used for three-dimensional given an alternative method for computing pairwise alignments more efficiently (such as using image features or landmarks).

@@

\subsection*{{\em Drosophila} experiments}
%
Oregon-R was used as wild type {\em Drosophila} strains. 
%
Embryos were collected and fixed at 22$^\circ$C. 
%
Monoclonal rabbit anti-dpERK (1:100, Cell signaling) and rat anti-Twist (1:500, a gift from Eric Wieschaus) were used to stain proteins of interest. 
%
DAPI (1:10,000, Vector Laboratories) was used to visualize nuclei, and Alexa Fluors (1:500, Invitrogen) were used as secondary antibodies. 
%
Histone-RFP strain was used to obtain time-lapse movie of gastrulating embryos at 22$^\circ$C. 
%
Embryos were loaded to the microfluidic device with PBST to keep them oxidized.

\subsection*{{\em Drosophila} microscopy}
%
Nikon A1-RS scanning confocal microscope, and the Nikon 60x Plan-Apo oil objective was used to image embryos. 
%
Embryos were collected, stained, and imaged together under the same microscope setting. 
%
End-on imaging was performed by using the microfluidics device described previously \citep{chung2010microfluidic}.
%
Images were collected at the focal plane $\sim$90~$\mathrm{\mu m}$ from the posterior pole of an embryo (see \fig~\ref{fig:DV_view}).

@@

\subsection*{Software and imaging data}
%
All algorithms and analysis were implemented in MATLAB\textsuperscript{\textregistered} (R2013b, The MathWorks, Natick, Massachusetts).
%
Software, including documentation and tutorials, along with the full imaging data sets used in this paper are available at \texttt{genomics.princeton.edu/stas/publications.html} under ``Codes and Data''.
